
---
title: "Cost-Effectiveness Analysis using a Markov Model in R"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document demonstrates how to perform a **cost-effectiveness analysis (CEA)** using a **Markov model** in R. We will:

1. Define the model parameters.
2. Create a Markov model with three health states: Healthy, Ill, and Death.
3. Perform deterministic sensitivity analysis (one-way).
4. Conduct probabilistic sensitivity analysis (Monte Carlo simulation).
5. Visualize results using a tornado diagram, cost-effectiveness plane, and acceptability curve.

---

## Step 1: Load Required Packages

```{r install-packages, message=FALSE}
# Install required packages if not already installed
if (!require("heemod")) install.packages("heemod", dependencies = TRUE)
if (!require("decisionSupport")) install.packages("decisionSupport", dependencies = TRUE)
if (!require("ggplot2")) install.packages("ggplot2", dependencies = TRUE)
if (!require("dplyr")) install.packages("dplyr", dependencies = TRUE)
```

---

## Step 2: Define Model Parameters

```{r define-parameters}
library(dplyr)

params <- list(
  population_size = 1000,
  p_healthy_to_ill = 0.10,  # Transition probability
  p_ill_to_death = 0.05,
  p_ill_to_healthy = 0.15,
  cost_healthy = 100,        # Cost per cycle
  cost_ill = 1000,
  cost_death = 0,
  utility_healthy = 1,       # QALY weights
  utility_ill = 0.5,
  utility_death = 0,
  cycles = 10,               # Time horizon
  discount_rate = 0.03
)
```

---

## Step 3: Build the Markov Model

```{r markov-model}
library(heemod)

# Define health states
state_healthy <- define_state(
  cost = params$cost_healthy,
  utility = params$utility_healthy
)

state_ill <- define_state(
  cost = params$cost_ill,
  utility = params$utility_ill
)

state_death <- define_state(
  cost = params$cost_death,
  utility = params$utility_death
)

# Define transition matrix
transition_matrix <- define_transition(
  state_names = c("Healthy", "Ill", "Death"),
  matrix = rbind(
    c(0.85, 0.10, 0.05),
    c(0.15, 0.80, 0.05),
    c(0.00, 0.00, 1.00)
  )
)

# Define and run the model
model <- define_model(
  transition = transition_matrix,
  Healthy = state_healthy,
  Ill = state_ill,
  Death = state_death
)

result <- run_model(
  model = model,
  cycles = params$cycles,
  cost = cost,
  effect = utility,
  init = c(Healthy = 1, Ill = 0, Death = 0),
  method = "half-cycle"
)
print(result)
```

---

## Step 4: Deterministic Sensitivity Analysis (One-Way)

```{r tornado-diagram}
sensitivity_params <- data.frame(
  parameter = c("p_healthy_to_ill", "p_ill_to_death", "cost_ill"),
  low = c(0.05, 0.02, 800),
  high = c(0.20, 0.10, 1200)
)

run_sensitivity <- function(param_name, value) {
  params[[param_name]] <- value
  result <- run_model(model, cycles = params$cycles)
  return(result$cost)
}

sensitivity_results <- apply(sensitivity_params, 1, function(row) {
  low <- run_sensitivity(row["parameter"], row["low"])
  high <- run_sensitivity(row["parameter"], row["high"])
  c(low = low, high = high)
})

sensitivity_df <- data.frame(
  Parameter = sensitivity_params$parameter,
  Low = sensitivity_results[1, ],
  High = sensitivity_results[2, ]
)

library(ggplot2)
ggplot(sensitivity_df, aes(x = reorder(Parameter, High - Low), y = High - Low)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Tornado Diagram", x = "Parameter", y = "Impact on Cost")
```

---

## Step 5: Probabilistic Sensitivity Analysis (PSA)

```{r psa}
library(decisionSupport)

input_estimates <- data.frame(
  variable = c("p_healthy_to_ill", "p_ill_to_death", "cost_ill"),
  distribution = c("beta", "beta", "normal"),
  lower = c(0.05, 0.02, 800),
  upper = c(0.20, 0.10, 1200)
)

mc_simulation <- mcSimulation(
  estimate = input_estimates,
  model_function = function(x) run_model(model, cycles = params$cycles)$cost,
  numberOfModelRuns = 1000
)

plot(mc_simulation, type = "cost-effectiveness-plane")
plot(mc_simulation, type = "ceac")
```

---

## Step 6: Summary Dashboard

```{r summary-dashboard}
summary <- data.frame(
  Health_State = c("Healthy", "Ill", "Death"),
  Average_Cost = result$cost,
  Average_Utility = result$effect
)

print(summary)

ggplot(summary, aes(x = Average_Utility, y = Average_Cost, label = Health_State)) +
  geom_point() +
  geom_text(vjust = -1) +
  labs(title = "Cost vs Utility", x = "Utility (QALY)", y = "Cost")
```

---

## Conclusion

This document shows how to perform cost-effectiveness analysis using a Markov model in R. We explored:

- Defining parameters for the model
- Building a Markov model with three health states
- Performing deterministic and probabilistic sensitivity analysis
- Visualizing results with tornado diagrams, cost-effectiveness planes, and utility-cost graphs
